---
title: "Visualizing Data"
author: 'Rafik Margaryan, MD, PhD'
date: 2025-03-17
institute: Ospedale Del Cuore Fondazione Monasterio
format: revealjs
execute:
  eval: true
  echo: true
  warning: false
  output: true
  include: true
transition: fade
show-slide-number: all
---

# Barplot in ggplot2

## Simple example

```{r}
library(ggplot2)
votes <- read.csv("data/votes_gg.csv")

ggplot()

```

Nothing happens!

## Feeding data to ggplot function

```{r eval=FALSE}
votes <- read.csv("data/votes_gg.csv")

ggplot(votes) + 
  geom_col()

```

Produces an error: `! Errore in geom_col(): geom_col() requires the following missing aesthetics:   x and y.`

## Mapping variables

```{r }
# Add x and y aesthetics

votes <- read.csv("data/votes_gg.csv")
ggplot(votes, aes(x = candidate, y = votes)) +
  geom_col()
```

And voila!

## Reshaping y limits

```{r}
ggplot(votes, aes(x = candidate, y = votes)) +
  geom_col() + scale_y_continuous(limits = c(0, 250))
```

## Renaming labs

```{r}
votes <- read.csv("data/votes_gg.csv")
ggplot(votes, aes(x = candidate, y = votes)) +
  geom_col() + scale_y_continuous(limits = c(0, 250)) +
  labs( x = "Candidate", y = "Votes", title = "Election Results")
```

## Add fill aesthetic mapping for geom_col

```{r}
votes <- read.csv("data/votes_gg.csv")
ggplot(votes, aes(x = candidate, y = votes)) +
  geom_col(aes(fill = candidate)) +
  scale_y_continuous(limits = c(0, 250)) + labs(x = "Candidate",y = "Votes", title = "Election Results")

```

## Using viridis scale to design for color blindness

```{r}
votes <- read.csv("data/votes_gg.csv")
ggplot(votes, aes(x = candidate, y = votes)) +
  geom_col(aes(fill = candidate)) +
  scale_fill_viridis_d("Candidate") +
  scale_y_continuous(limits = c(0, 250)) +labs(x = "Candidate",y = "Votes",title = "Election Results")
```

## Adjust ggplot theme

```{r}
votes <- read.csv("data/votes_gg.csv")
ggplot(votes, aes(x = candidate, y = votes)) +
  geom_col(aes(fill = candidate)) +
  scale_fill_viridis_d("Candidate") +
  scale_y_continuous(limits = c(0, 250)) +
  labs(x = "Candidate",y = "Votes", title = "Election Results") + theme_classic()
```

## Saving file

```{r}
# Save file

votes <- read.csv("data/votes_gg.csv")

p <- ggplot(votes, aes(x = candidate, y = votes)) +
  geom_col(aes(fill = candidate)) +
  scale_fill_viridis_d("Candidate") +
  scale_y_continuous(limits = c(0, 250)) +
  labs(
    x = "Candidate",
    y = "Votes",
    title = "Election Results"
  ) +
  theme_classic()

ggsave(
  "images/votes.png",
  plot = p,
  width = 1200,
  height = 900,
  units = "px"
)
print('Done...!')
```

# geom_point in ggplot2

## simple example

```{r}
load("data/candy.RData")
ggplot(
  candy,
  aes(x = price_percentile, y = sugar_percentile)
) + geom_point()
```

Data is from `candy data`, [^longnote]. 


[^longnote]: data is from Tomic et al. (2010), measurements from sensory analysis (professional tasting) on a number of candy products obtained by sensory labs.

## Adding theme 

```{r}
ggplot(
  candy,
  aes(x = price_percentile, y = sugar_percentile)
) +   geom_point() +   labs(
    x = "Price",
    y = "Sugar",
    title = "Price and Sugar"
  ) +
  theme_classic()
```

## Introduce size and color aesthetic


```{r}
ggplot(
  candy,
  aes(x = price_percentile, y = sugar_percentile)
) +geom_jitter( color = "darkorchid", size = 2 ) + labs( x = "Price", y = "Sugar",
    title = "Price and Sugar" ) + theme_classic()
```

## Introduce point shape and fill color

```{r}
ggplot(candy,
  aes(x = price_percentile, y = sugar_percentile)) +
  geom_jitter(color = "darkorchid",
    fill = "orchid",
    shape = 21,
    size = 2) +
  labs(x = "Price",y = "Sugar",
    title = "Price and Sugar"
  ) + theme_classic()
```

# geom_line

## Basic Line Plot

```{r}
load("data/anita.RData")
ggplot(anita, aes(x = timestamp, y = wind)) +
  geom_point()
```
Data [^longnote]. 


[^longnote]: generated using R.

## Add line

```{r}
ggplot(anita, aes(x = timestamp, y = wind)) +
  geom_line()
```

## Combine both geoms

```{r}
ggplot(anita, aes(x = timestamp, y = wind)) +
  geom_line() +
  geom_point(color = "deepskyblue4")
```

## Experiment with geom_line and geom_point aesthetics

```{r}
ggplot(anita, aes(x = timestamp, y = wind)) +
  geom_line(
    linetype = 1,
    linewidth = 0.5) +
  geom_point( color = "deepskyblue4", size = 2)
```

## Add labels and adjust theme

```{r}
ggplot(anita, aes(x = timestamp, y = wind)) +
  geom_line(linetype = 1,linewidth = 0.5) +
  geom_point(color = "deepskyblue4",size = 2) +
  labs(y = "Wind Speed (Knots)", x = "Date", title = "Hurricane Anita" ) + theme_classic()
```

## Add horizontal line to demarcate hurricane status

```{r}
ggplot(anita, aes(x = timestamp, y = wind)) + geom_line( linetype = 1, linewidth = 0.5 ) + geom_point( color = "deepskyblue4",size = 2) +
  geom_hline( linetype = 3,yintercept = 64) +
  labs(y = "Wind Speed (Knots)",x = "Date", title = "Hurricane Anita"
  ) + theme_classic()
```
